---
title: "IAS poster visuals"
author: "Karishma Srikanth"
date: "`r Sys.Date()`"
output: html_document
---

## Overview

### Background

Write a little bit about XX community based program


### Data Description

**The response variable is:** 

*Re-engaged to treatment Value (rtt_binary):* 1 = the patient traced and returned to treatment,0= the patient traced but not found or found but not returned to treatment.

**The predicting variables are:** 


*gender:* The patient's gender of patient(Male , Female).
*age bands:* Age group.
*on ART duration:.
*duration between Missed date and Tracing date:.
*healthfac_type:* : Hospital, Primary health care and others
*tracing_method_int:* Indicates patients tracing method.

***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*** Check to see if necessary packages are installed.  If not, install them ***
```{r include=FALSE}
# lists all packages needed in this notebook
packages_list = list('knitr', 'tidyverse', 'corrplot', 'caTools', 'car',
                     'glmnet','caret','kknn','rpart','rpart.plot','randomForest',"pROC","Metrics","scales","e1071","magrittr","partykit","dplyr","magrittr","party","MLmetrics","mlr","Metrics")


# checks to see if package is installed.  If not, installs and loads it
for (i in packages_list) {
  if (!require(i, character.only=TRUE)) install.packages(i, character.only=TRUE); library(i, character.only=TRUE)
}

# removes the assignment of i and packages_list
rm(i, packages_list)
```


## Preparing the Data

### Reading data

```{r, message=FALSE}
# Read the data using read_csv
dat <- read.csv("eth_ITT.csv",header=TRUE) #KS: use output from dataout and rename eth_IIT
```

```{r}
# check variables 
#variable.names(dat)
dim(dat)
```


```{r}
# Keep only variables useful for our analysis 
df<- dat[c(2,4,6,7,8,9,12,16,28,29,31,37,39, 41, 42, 44)]
#variable.names(df) # view the variables selected 
```

### Data Prep

```{r}
# Convert categorical variables to factors
df$lip= as.factor(df$lip)
df$zone= as.factor(df$zone)
df$gender= as.factor(df$gender)
df$age_bands= as.factor(df$age_bands)

df$rtt_binary= as.factor(df$rtt_binary)
df$year_art_init_gap = df$year_art_init_gap
df$missed_ltfu_dur_grp= as.factor(df$missed_ltfu_dur_grp)
df$missed_custom_report_dur_grp= as.factor(df$missed_custom_report_dur_grp)
df$healthfac_type= as.factor(df$healthfac_type)
df$fiscal_year= as.factor(df$fiscal_year)
df$ltfu_tracing_method_1= as.factor(df$ltfu_tracing_method_1)

```

```{r}
# summarize sample size by fiscal year
summary_df<- df %>%
  group_by(fiscal_year) %>%
  summarise( count=n()
    
  )
#summary_df

```


```{r}
#Mutate tracing method_one
df<- df %>%
  mutate(tracing_method_int= case_when (ltfu_tracing_method_1 %in% c("other","informed_by_family","informed_by_treatment_sup")~"other_methods",
       ltfu_tracing_method_1 %in% c("phone") ~ "phone" , 
       ltfu_tracing_method_1 %in% c("home_visit" ) ~ "home_visit"))

 df$tracing_method_int = as.factor(df$tracing_method_int) # convert it to a factor
 #variable.names(df)
```

```{r}
# Creating new variables

# Duration on ART before interrupted treatment 
df$art_int_date <- as.Date(df$client_art_start_date)
df$itt_date <- as.Date(df$missed_appointment_date)
df$on_art_duration <- difftime(df$itt_date,df$art_int_date,units = "days")

df <- df %>%
  mutate(duration_art_missgrp = case_when(on_art_duration < 30 ~ "< 30 days",
                                          on_art_duration >= 30 & on_art_duration < 90 ~ "30-90 days",
                                           on_art_duration >= 90 & on_art_duration < 365 ~ "90-365 days",
                                           on_art_duration >= 365 & on_art_duration <= 1095 ~ "1-3 yrs",
                                           on_art_duration > 1095 ~ "3+ yrs"))
# convert it as factor
df$duration_art_missgrp = as.factor(df$duration_art_missgrp)
```

```{r}
# Re-create age_bands
df <- df %>%
   mutate(age_bands2 = case_when(
    age %in% c(1:14) ~ "<15",
    age %in% c(15:19) ~ "15-19",
    age %in% c(20:29) ~ "20-29",
    age %in% c(30:39) ~ "30-39",
    age %in% c(40:49) ~ "40-49",
    age %in% c(50:100) ~ "50+"))
 df$age_bands2 = as.factor(df$age_bands2) # convert it to a factor
```

```{r}
#Number of days between missed appt date and tracing date initiated 
df$tracing_date_int <- as.Date(df$ltfu_tracing_date)
df$missed_tracing_duration <- difftime(df$tracing_date_int,df$itt_date,units = "days")
df <- df %>%
  mutate(missed_tracing_dategrp = case_when(missed_tracing_duration < 30 ~ "< 30 days",
                                         missed_tracing_duration  >= 30 & missed_tracing_duration  < 60 ~ "30-60 days",
                                          missed_tracing_duration  >= 60 & missed_tracing_duration  <= 90 ~ "60-90 days",
                                          missed_tracing_duration  > 90 ~ "90+ days"))

df$duration_art_missgrp = as.factor(df$duration_art_missgrp)
df$missed_tracing_dategrp = as.factor(df$missed_tracing_dategrp)
#variable.names(df)
```


```{r}
#Reduce sample size to 2023

df_2023<- df %>%
  filter(fiscal_year=="2023") 

# Dropping observations with missing values NA's (from 1761 to 1535 due to missing ART initiation date )

df_2023_clean <- df_2023 %>% drop_na()
# Check  NA -missing values for each column  
#colSums(is.na(df_2023_clean))

df_2023_clean_tbl <- df_2023_clean %>% 
    rename_with(~ str_to_title(.), everything()) %>%
    rename_with(~ str_replace(., "_", " "), everything()) %>% 
  mutate(`Duration art_missgrp` = fct_relevel(`Duration art_missgrp`, c("< 30 days", "30-90 days",
                                                                        "90-365 days", "1-3 yrs", "3+ yrs"))) %>% 
  rename(`Type of Health Facility` = `Healthfac type`,
         `Duration between ART initiation date and treatment interruption date` = `Duration art_missgrp`,
         `Duration between tracing date and treatment interruption date` = `Missed tracing_dategrp`,
         `Tracing Method` = `Tracing method_int`) %>% 
#mutate(`Rtt binary` = as.numeric(`Rtt binary`)) %>% 
  mutate(rtt = ifelse(`Rtt binary` == 0, "Not returned to treatment", "Returned to treatment")) %>% 
  mutate(`Tracing Method` = str_replace_all(`Tracing Method`, "_", " ") %>% str_to_title())


table1::table1(~ Gender + `Age bands` + `Type of Health Facility` + 
                 `Duration between ART initiation date and treatment interruption date` +
                 `Duration between tracing date and treatment interruption date` + `Tracing Method` | rtt, data = df_2023_clean_tbl,
               #overall = F,
               #extra.col=list(`P-value`=pvalue),#add in p-values
               topclass = "Rtable1-grid Rtable1-shade Rtable1-times", 
               caption = "Table 1: Baseline Cohort Demographics") 
```

```{r}
#Dropping columns not considered further in the analysis
df_2023_rtt<- df_2023_clean[, -c(1,2,3,5, 6,7,8,9,10,12,13,15,17,18,19,22,23)]
variable.names(df_2023_rtt)
```